---
title: "Statistical Inference Project Part2"
author: "GaryFH"
date: "April 26, 2017"
output:
  pdf_document: default
  html_document: default
---
#       OVERVIEW PART 2

###     Basic Inferential Data Analysis of "ToothGrowth data" 

####  In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package. 

##### Setup work environment
```{r echo=FALSE}
        library(knitr)
        library(ggplot2)
        library(dplyr)
        library(tidyr)
```


###First load data

```{r load}
        data("ToothGrowth")
        dtooth<-ToothGrowth
        dftooth<-tbl_df(dtooth)
```

###Overview/summary of "ToothGrowth"

```{r overview}
        summary(dftooth)
        dftooth
```

### Find mean of supply mode:

```{r findmeans}
        df2<-group_by(dtooth,supp)
        df3<-df3<-summarize(df2,length=mean(len))
        df3
```

### Plot 3 - Tooth size vrs kind of supplement used

```{r, Length-supplement}

        g3<-ggplot(dftooth, aes(x=supp,y=len))
        plot3<-g3+geom_boxplot(aes(fill=supp))+labs(x="Type of Supplement", y="Length of Tooth", title="Length of Tooth vrs supplement type")
        plot3
```

### Plot 4 - Tooth size vrs Vitamin Dose

```{r vitaminDose}
        dftooth$dose<-as.factor(dftooth$dose)
        g4<-ggplot(dftooth, aes(x=dose,y=len))
        plot4<-g4+geom_boxplot(aes(fill=dose))+labs(x="Dose-milligrams", y="Length of Tooth", title="Length of Tooth vrs Dose Strength")
        plot4
```



### Use confidence intervals to compare tooth growth by supp.

```{r CI_supp}

        t.test(dftooth$len[dftooth$supp=="OJ"], dftooth$len[dftooth$supp=="VC"], paired = FALSE, var.equal = FALSE)$conf.int
```

#### The above confidence interval contains 0 therefore we can conclude that the supplement type has no impact on tooth growth


### Use confidence intervals to compare tooth growth by dose.

```{r CI_dose}

        t.test(dftooth$len[dftooth$dose==2], dftooth$len[dftooth$dose==1], paired = FALSE, var.equal = FALSE)$conf.int
```

#### The above confience interval does not contain 0 therefore we can conclude that the supplement type does impact on tooth growth - the higher dose has higher growth rate.